---
title: "TLSA_DataExploration"
author: "Yu-Hsuan Chou"
date: "2021/8/6"
output:
  html_document: default
  pdf_document: default
refine: Hsiao, 0923
---

```{r}
library(data.table)
library(ggplot2)
library(fixest)
library(ggfortify)
library(stargazer)
library(stringr)
```


```{r}
subject.dt <- as.data.table(read.csv("/Users/hsiao/Desktop/Projects/HRS_Chao/LTC_TSLA/main_subject_1026.csv"))
fmlymmb.dt <- as.data.table(read.csv("/Users/hsiao/Desktop/Projects/HRS_Chao/LTC_TSLA/family_member_1010.csv"))

```

```{r}
# what happened on augmented spouses
# Chou miscoded. She used the qser_no for all subjects over multiple observed years
# After change of the code, 405 problematic remains, while 326 of them is NAs, therefore can be omited.
# NA are mainnly other relationship in 1996 which should be omitted earlier
# the 79 remaining is mainly consisted of 同居老伴 ( 1993, 1996 old 沒有問同居)
# 
# spou<- fmlymmb.dt[relationship ==2 & ObsPartnerY ==0]
# sum(is.na(spou$marstat)) # 326
# spou<- spou[! is.na(marstat)]
# who <- fmlymmb.dt[ ObsPartnerY ==0 & relationship ==2 & is.na(marstat)]
# table(who)
```

```{r}
# duplicated people
# main change :
# detected_1 added to detect if one fam member only exist in specific year, is he the only one ? 
# otherwise he will go through k_grouping.
# unif stands for the unificated code

# test<- copy(fmlymmb.dt)
# sum(duplicated(test$unif))
```

# Family Member Personalities

### Preprocess
```{r}
fmlymmb.dt$male = ifelse(fmlymmb.dt$male == 1, 1, 0)  # male 0 = females
fmlymmb.dt$partner = ifelse(fmlymmb.dt$told_relation == 2, 1, 0) 
fmlymmb.dt$working = ifelse(fmlymmb.dt$work == 1, 1, 0)
fmlymmb.dt$unmarried = ifelse(fmlymmb.dt$marstat == 6, 1, 0) 
fmlymmb.dt$attend_at_least_uni = ifelse(fmlymmb.dt$edu >= 13, 1, 0) 
fmlymmb.dt$attend_at_least_high_school = ifelse(fmlymmb.dt$edu >=10, 1, 0) 
# age above 15 & unmarry -> 6, if under 15 -> NA
fmlymmb.dt$livTogether = ifelse(fmlymmb.dt$livstat == 0, 1, 0)
fmlymmb.dt$primaryADL = ifelse(fmlymmb.dt$adl_who_help == fmlymmb.dt$told_relation, 1, 0)
fmlymmb.dt$primaryIADL = ifelse(fmlymmb.dt$iadl_who_help == 
                                  fmlymmb.dt$told_relation, 1, 0)
fmlymmb.dt$fmNeedADL = ifelse(! is.na(fmlymmb.dt$adl_who_help) , 1, 0)
fmlymmb.dt$fmNeedIADL = ifelse(! is.na(fmlymmb.dt$iadl_who_help) , 1, 0)
fmlymmb.dt$subject_residency = as.character(fmlymmb.dt$subject_residency)
# Define degree of intimacy
```

```{r}
subject.dt[, y_qser_no := paste0( survey_year, qser_no)]
fmlymmb.dt[, y_qser_no := paste0(survey_year, qser_no)]

non_married.obs <- subject.dt[marstat == 0]$y_qser_no
fmlymmb.dt.nmar <- fmlymmb.dt[y_qser_no %in% non_married.obs]
fmlymmb.dt$ObsPartnerY <- ifelse(fmlymmb.dt$y_qser_no %in% non_married.obs, 0, 1)

# fmlymmb.dt<- fmlymmb.dt[! (ObsPartnerY ==0 & relationship ==2 & is.na(marstat))]
```

### Merging with Subject Data

```{r}
subject.dt$rsex = ifelse(subject.dt$rsex == 1, 1, 0)
fmlymmb.dt <- merge(fmlymmb.dt, 
                    subject.dt[,c("qser_no", "rsex", "survey_year", "rage")], 
                    all.x = TRUE, 
                    by = c("qser_no", "survey_year"))
```


### Summary
```{r}

fm.with.need <- fmlymmb.dt[fmNeedADL == 1]
fm.primary.c <- fmlymmb.dt[primaryADL == 1]
fm.whot.need <- fmlymmb.dt[fmNeedADL == 0]
print(paste("People in families with need:", nrow(fm.with.need)))

print(paste("Number of Primary Caregiver:", nrow(fm.primary.c)))

print(paste("People in families without need:", nrow(fm.whot.need)))
```

#### Age, male, Education, Marriage Status
```{r}
rbind(fm.with.need[ObsPartnerY == 1,.(AGE = mean(age, na.rm = TRUE), 
                      EDU = mean(edu, na.rm = TRUE), 
                      SEX = mean(male, na.rm = TRUE),
                      PARTNER = mean(partner, na.rm = TRUE),
                      WORK = mean(working, na.rm = TRUE),
                      UNMARRIED = mean(unmarried, na.rm = TRUE),
                      LVTOGETER = mean(livTogether, na.rm = TRUE),
                      SUBSEX = mean(rsex, na.rm = TRUE), 
                      SUBAGE = mean(rage, na.rm = TRUE)
                      )],
      fm.with.need[ObsPartnerY == 0,.(AGE = mean(age, na.rm = TRUE), 
                      EDU = mean(edu, na.rm = TRUE), 
                      SEX = mean(male, na.rm = TRUE),
                      PARTNER = mean(partner, na.rm = TRUE),
                      WORK = mean(working, na.rm = TRUE),
                      UNMARRIED = mean(unmarried, na.rm = TRUE),
                      LVTOGETER = mean(livTogether, na.rm = TRUE),
                      SUBSEX = mean(rsex, na.rm = TRUE), 
                      SUBAGE = mean(rage, na.rm = TRUE)
                      )],
      fm.whot.need[ObsPartnerY == 1,.(AGE = mean(age, na.rm = TRUE), 
                      EDU = mean(edu, na.rm = TRUE), 
                      SEX = mean(male, na.rm = TRUE),
                      PARTNER = mean(partner, na.rm = TRUE),
                      WORK = mean(working, na.rm = TRUE),
                      UNMARRIED = mean(unmarried, na.rm = TRUE),
                      LVTOGETER = mean(livTogether, na.rm = TRUE),
                      SUBSEX = mean(rsex, na.rm = TRUE), 
                      SUBAGE = mean(rage, na.rm = TRUE)
                      )],
      fm.whot.need[ObsPartnerY == 0,.(AGE = mean(age, na.rm = TRUE), 
                      EDU = mean(edu, na.rm = TRUE), 
                      SEX = mean(male, na.rm = TRUE),
                      PARTNER = mean(partner, na.rm = TRUE),
                      WORK = mean(working, na.rm = TRUE),
                      UNMARRIED = mean(unmarried, na.rm = TRUE),
                      LVTOGETER = mean(livTogether, na.rm = TRUE),
                      SUBSEX = mean(rsex, na.rm = TRUE), 
                      SUBAGE = mean(rage, na.rm = TRUE)
                      )],
      fm.primary.c[ObsPartnerY == 1,.(AGE = mean(age, na.rm = TRUE), 
                      EDU = mean(edu, na.rm = TRUE), 
                      SEX = mean(male, na.rm = TRUE),
                      PARTNER = mean(partner, na.rm = TRUE),
                      WORK = mean(working, na.rm = TRUE),
                      UNMARRIED = mean(unmarried, na.rm = TRUE),
                      LVTOGETER = mean(livTogether, na.rm = TRUE),
                      SUBSEX = mean(rsex, na.rm = TRUE), 
                      SUBAGE = mean(rage, na.rm = TRUE)
                      )],
      fm.primary.c[ObsPartnerY == 0,.(AGE = mean(age, na.rm = TRUE), 
                      EDU = mean(edu, na.rm = TRUE), 
                      SEX = mean(male, na.rm = TRUE),
                      PARTNER = mean(partner, na.rm = TRUE),
                      WORK = mean(working, na.rm = TRUE),
                      UNMARRIED = mean(unmarried, na.rm = TRUE),
                      LVTOGETER = mean(livTogether, na.rm = TRUE),
                      SUBSEX = mean(rsex, na.rm = TRUE), 
                      SUBAGE = mean(rage, na.rm = TRUE)
                      )])
```

### Regression
#### Families with need
```{r}
fm.with.need$SubSameSex <- ifelse(fm.with.need$male == fm.with.need$rsex, 1, 0)
need.reg1 <- feols(primaryADL ~ male| survey_year, fm.with.need)
need.reg2 <- feols(primaryADL ~ male + age + age**2 | survey_year, fm.with.need)
need.reg3 <- feols(primaryADL ~ male + age + age**2 + edu | survey_year, fm.with.need)
need.reg4 <- feols(primaryADL ~ male + age + age**2 + edu + working | survey_year, fm.with.need)
need.reg5 <- feols(primaryADL ~ male + age + age**2 + edu + working + unmarried | survey_year, fm.with.need)
need.reg6 <- feols(primaryADL ~ male + age + age**2 + edu + working + unmarried + SubSameSex | survey_year, fm.with.need)
need.reg7 <- feols(primaryADL ~ male + age + age**2 + edu + working + unmarried + SubSameSex + partner | survey_year, fm.with.need)
need.reg8 <- feols(primaryADL ~ male + age + age**2 + edu + working + unmarried + SubSameSex + partner + livTogether | survey_year, fm.with.need)
need.reg9 <- feols(primaryADL ~ male + age + age**2 + edu + working + unmarried + SubSameSex + partner + livTogether + subject_in_nursing_house | survey_year, fm.with.need)
need.reg10 <- feols(primaryADL ~ male + age + age**2 + edu + working + unmarried + SubSameSex + partner + livTogether  | survey_year, fm.with.need[ subject_in_nursing_house ==0])
etable(need.reg1, need.reg2, need.reg3, need.reg4, need.reg5, need.reg6, need.reg7, need.reg7, need.reg8, need.reg10)


```

### Decision Tree
```{r}
library(rpart)
library(rpart.plot)
```

```{r}
simple.tree <- rpart(primaryADL ~ male + age + age**2 + 
                       edu + partner + working + SubSameSex + 
                       unmarried + livTogether, data = fm.with.need, cp=.008)
rpart.plot(simple.tree, faclen=0, 
           box.palette="RdBu", 
           fallen.leaves=TRUE, 
           shadow.col="gray", 
           nn=TRUE)
```

```{r}
simple.tree <- rpart(primaryADL ~ male + age + age**2 + 
                       edu + partner + working + SubSameSex + 
                       unmarried, data = fm.with.need, cp=.003)
rpart.plot(simple.tree, faclen=0, 
           box.palette="RdBu", 
           fallen.leaves=TRUE, 
           shadow.col="gray", 
           nn=TRUE)
```

```{r}
simple.tree <- rpart(primaryADL ~ male + age + age**2 + 
                       edu + working + SubSameSex + 
                       unmarried, data = fm.with.need, cp=.003)
rpart.plot(simple.tree, faclen=0, 
           box.palette="RdBu", 
           fallen.leaves=TRUE, 
           shadow.col="gray", 
           nn=TRUE)
```

```{r}
# which spouse would be the main giver
## omitted some very important variables : partner income, partner health status
simple.tree <- rpart(primaryADL ~ male + age + age**2 + 
                       edu + working + SubSameSex + 
                       unmarried, data = fm.with.need[ partner ==1 ], cp=.003)
rpart.plot(simple.tree, faclen=0, 
           box.palette="RdBu", 
           fallen.leaves=TRUE, 
           shadow.col="gray", 
           nn=TRUE)
```

### PCA Clustering
```{r}
data.pca <- fm.with.need[,c("male", "age", "edu", "working", "unmarried", "SubSameSex", "partner", "primaryADL")]
data.pca$primaryADL <- factor(data.pca$primaryADL)
simple.pca <- prcomp(formula = ~ .,  
              data = data.pca[,1:7], 
              scale = TRUE)
# take away livtogether

```

```{r}
plot(simple.pca, type="line") 
abline(h=1, col="blue")
```

```{r}
simple.pca$rotation[,1:3]
first.pca <- simple.pca$rotation[, 1]   #  第一主成份
second.pca <- simple.pca$rotation[, 2]  #  第二主成份
third.pca <- simple.pca$rotation[, 3]   #  第三主成份
dotchart(first.pca[order(first.pca, decreasing=FALSE)] ,   # 排序後的係數
         main="Loading Plot for PC1",                      # 主標題
         xlab="Variable Loadings",                         # x軸的標題
         col="black")

dotchart(second.pca[order(second.pca, decreasing=FALSE)] ,   # 排序後的係數
         main="Loading Plot for PC2",                      # 主標題
         xlab="Variable Loadings",                         # x軸的標題
         col="black")

dotchart(third.pca[order(third.pca, decreasing=FALSE)] ,   # 排序後的係數
         main="Loading Plot for PC3",                      # 主標題
         xlab="Variable Loadings",                         # x軸的標題
         col="black")
```

```{r}
library(devtools)
install_github('sinhrks/ggfortify')
library(ggfortify); library(ggplot2)
autoplot(simple.pca, data = data.pca, colour = 'primaryADL', alpha=0.1, loadings = TRUE, loadings.colour = 'gray', loadings.label = TRUE, loadings.label.alpha = 0.4, loadings.label.colour = "black")
```


## Excluding Married Observations

### Summary
#### Clean Those with Partners
```{r}
fmlymmb.dt.nmar <- fmlymmb.dt[ObsPartnerY == 0]
odd.obs <- unique(fmlymmb.dt.nmar[told_relation==2]$qser_no) # 同居但沒結婚 72 
fmlymmb.dt.nmar <- fmlymmb.dt.nmar[!(fmlymmb.dt.nmar$qser_no %in% odd.obs)]

table(fmlymmb.dt.nmar$told_relation)

# fmlymmb.dt.nmar<- fmlymmb.dt.nmar[ told_relation %in% c()]
```

```{r}
fm.with.need <- fmlymmb.dt.nmar[fmNeedADL == 1]
fm.primary.c <- fmlymmb.dt.nmar[primaryADL == 1]
fm.whot.need <- fmlymmb.dt.nmar[fmNeedADL == 0]
print(paste("People in families with need:", nrow(fm.with.need)))
print(paste("Number of Primary Caregiver:", nrow(fm.primary.c)))
print(paste("People in families without need:", nrow(fm.whot.need)))
```

#### Age, male, Education, Marriage Status
```{r}
rbind(fm.with.need[,.(AGE = mean(age, na.rm = TRUE), 
                      EDU = mean(edu, na.rm = TRUE), 
                      SEX = mean(male, na.rm = TRUE),
                      PARTNER = mean(partner, na.rm = TRUE),
                      WORK = mean(working, na.rm = TRUE),
                      UNMARRIED = mean(unmarried, na.rm = TRUE),
                      LVTOGETER = mean(livTogether, na.rm = TRUE),
                      SUBSEX = mean(rsex, na.rm = TRUE), 
                      SUBAGE = mean(rage, na.rm = TRUE),
                      Sub_in_nursing = mean( subject_in_nursing_house, na.rm = TRUE)
                      )],
      fm.whot.need[,.(AGE = mean(age, na.rm = TRUE), 
                      EDU = mean(edu, na.rm = TRUE), 
                      SEX = mean(male, na.rm = TRUE),
                      PARTNER = mean(partner, na.rm = TRUE),
                      WORK = mean(working, na.rm = TRUE),
                      UNMARRIED = mean(unmarried, na.rm = TRUE),
                      LVTOGETER = mean(livTogether, na.rm = TRUE),
                      SUBSEX = mean(rsex, na.rm = TRUE), 
                      SUBAGE = mean(rage, na.rm = TRUE),
                      Sub_in_nursing = mean( subject_in_nursing_house, na.rm = TRUE)
                      )],
      fm.primary.c[,.(AGE = mean(age, na.rm = TRUE), 
                      EDU = mean(edu, na.rm = TRUE), 
                      SEX = mean(male, na.rm = TRUE),
                      PARTNER = mean(partner, na.rm = TRUE),
                      WORK = mean(working, na.rm = TRUE),
                      UNMARRIED = mean(unmarried, na.rm = TRUE),
                      LVTOGETER = mean(livTogether, na.rm = TRUE),
                      SUBSEX = mean(rsex, na.rm = TRUE), 
                      SUBAGE = mean(rage, na.rm = TRUE),
                      Sub_in_nursing = mean( subject_in_nursing_house, na.rm = TRUE)
                      )])
```

### Regression
#### Family with Need (Observations Not Having Partners)
```{r}
fm.with.need$SubSameSex <- ifelse(fm.with.need$male == fm.with.need$rsex, 1, 0)
need.reg1 <- feols(primaryADL ~ male| survey_year, fm.with.need)
need.reg2 <- feols(primaryADL ~ male + age + age**2 | survey_year, fm.with.need)
need.reg3 <- feols(primaryADL ~ male + age + age**2 + edu | survey_year, fm.with.need)
need.reg4 <- feols(primaryADL ~ male + age + age**2 + edu + working | survey_year, fm.with.need)
need.reg5 <- feols(primaryADL ~ male + age + age**2 + edu + working + unmarried | survey_year, fm.with.need)
need.reg6 <- feols(primaryADL ~ male + age + age**2 + edu + working + unmarried + SubSameSex | survey_year, fm.with.need)
need.reg7 <- feols(primaryADL ~ male + age + age**2 + edu + working + unmarried + SubSameSex + livTogether | survey_year, fm.with.need)
need.reg8 <- feols(primaryADL ~ male + age + age**2 + edu + working + unmarried + SubSameSex + livTogether | survey_year, fm.with.need)
need.reg9 <- feols(primaryADL ~ male + age + age**2 + edu + working + unmarried + SubSameSex + livTogether + subject_in_nursing_house + subject_residency | survey_year+ subject_residency, fm.with.need)
need.reg10 <- feols(primaryADL ~ male + age + age**2 + attend_at_least_high_school + working + unmarried + SubSameSex + livTogether + subject_in_nursing_house | survey_year+ subject_residency , fm.with.need)
need.reg11 <- feols(primaryADL ~ male + age + age**2 + attend_at_least_uni + working + unmarried + SubSameSex + livTogether + subject_in_nursing_house  | survey_year +subject_residency, fm.with.need)

etable(need.reg1, need.reg2, need.reg3, need.reg4, need.reg5, need.reg6, need.reg7, need.reg8, need.reg9, need.reg10, need.reg11)
```

### Decision Tree
```{r}
simple.tree <- rpart(primaryADL ~ male + age + age**2 + 
                       edu + working + 
                       unmarried + livTogether, data = fm.with.need, cp=.008)
```

```{r}
rpart.plot(simple.tree, faclen=0, 
           box.palette="RdBu", 
           fallen.leaves=TRUE, 
           shadow.col="gray", 
           nn=TRUE)
```

```{r}
simple.tree <- rpart(primaryADL ~ male + age + age**2 + 
                       edu + working + SubSameSex + 
                       unmarried, data = fm.with.need, cp=.003)
rpart.plot(simple.tree, faclen=0, 
           box.palette="RdBu", 
           fallen.leaves=TRUE, 
           shadow.col="gray", 
           nn=TRUE)
```
```{r}
simple.tree <- rpart(primaryADL ~ male + age + age**2 + 
                       attend_at_least_uni + attend_at_least_high_school + working + SubSameSex + 
                       unmarried, data = fm.with.need, cp=.003)
rpart.plot(simple.tree, faclen=0, 
           box.palette="RdBu", 
           fallen.leaves=TRUE, 
           shadow.col="gray", 
           nn=TRUE)


```


### PCA
```{r}
data.pca <- fm.with.need[,c("male", "age", "edu", "working", 
                            "unmarried", "SubSameSex", "primaryADL")]
data.pca$primaryADL <- factor(data.pca$primaryADL)
simple.pca <- prcomp(formula = ~ .,  
              data = data.pca[,1:6], 
              scale = TRUE)   
```

```{r}
plot(simple.pca, type="line") 
abline(h=1, col="blue")
```

```{r}
simple.pca$rotation[,1:3]
first.pca <- simple.pca$rotation[, 1]   #  第一主成份
second.pca <- simple.pca$rotation[, 2]  #  第二主成份
third.pca <- simple.pca$rotation[, 3]   #  第三主成份
dotchart(first.pca[order(first.pca, decreasing=FALSE)] ,   # 排序後的係數
         main="Loading Plot for PC1",                      # 主標題
         xlab="Variable Loadings",                         # x軸的標題
         col="black")

dotchart(second.pca[order(second.pca, decreasing=FALSE)] ,   # 排序後的係數
         main="Loading Plot for PC2",                      # 主標題
         xlab="Variable Loadings",                         # x軸的標題
         col="black")

dotchart(third.pca[order(third.pca, decreasing=FALSE)] ,   # 排序後的係數
         main="Loading Plot for PC3",                      # 主標題
         xlab="Variable Loadings",                         # x軸的標題
         col="black")
```

```{r}
# !!
autoplot(simple.pca, data = data.pca, colour = 'primaryADL', alpha=0.1, loadings = TRUE, loadings.colour = 'gray', loadings.label = TRUE, loadings.label.alpha = 0.4, loadings.label.colour = "black")
```